---
title: "Incarceration Trends in the United States (1970-2018)"
author: "Greta Carsten"
date: '2022-11-18'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/a4-helpers.R")
source("../source/analysis.R")
```

## Abstract 
Throughout this assignment, I looked very closely and analyzed **Incarceration Trends in the United States from the years 1970 to 2018**. Within this data, we can look at the rates of different *demographics* in jail or prison. By looking at this, we can find many trends of *inequality* within the **United States legal system** regarding *incarceration*. Is is extremely important that we continue to study this data in order for us to make **change** within the **United States**. 

# Test (TODO: Remove before submission!!)
```{r echo=FALSE}
q1_string <- test_query1()
num <- 5
```
As you know, this is a markdown report. The [R Markdown Cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf) is **useful!**

* Here's a test of a dynamic variable. Say **"`r q1_string`!!!** 
* Here's another test. A vector
of numbers: `r test_query2(num)`.

Recall that _{r echo=TRUE}_ can be used to display the code, 
which is very useful when developing and debugging a report. Here's an example: 

```{r echo=TRUE}
num <- 5
sum <- sum(test_query2(num))
print(sum)
```

## Introduction 

**Incarceration Trends within the United States** is something that we need to look analyze, and closely study in order to make *change* within the **United States legal system**. In the *justice system* of the **United States** many people are being wrongfully accused of a crime, and put in jail or prison. Throughout this data, we can look at the amount of people in jail by different *demographics* such as race, gender, county, state, land density and many more factors. Analyzing this data can show us trends, and see if there are specific *places* or *demographics* that have higher rates of incarceration, and we can compare these ratios and see if one place has higher amounts of incarceration levels. It is extremely important that we study this *incarceration trends* as they can help us make changes in our current legal system. Throughout this assignment, I closely analyzed this data, and learned various things. 


## Data Summary 

jail_pop_ratio <- us_total_jail_pop_county_state_year_2018 %>%
  select(
    jail_pop_ratio,
    state
  )
View(jail_pop_ratio)

This code is showing the ratio of the top ten most populated states, and a ratio 
of their total jail population compared to their total state population. From this data, we can see that Georgia, Florida, and Pennsylvania have the highest ratios of jail populations. From this data we can study why these states have a significantly higher ratio than the other most highly populated states in the United States. 

get_jail_pop_by_states <- function(states) {
  states_data <- incarceration_trends %>%
    filter(state %in% states) %>%
    group_by(year, state) %>%
    summarise(total_jail_pop = sum(total_jail_pop, na.rm = TRUE))
  return(states_data)
  
This function allows us to see how there are some states with much higher incarceration trends than others. For example, California has a much higher incarcerated population than Washington. Is this solely because of Californias higher population, more crime being committed in California, or are there other factors such as systemic racism. 



get_king_county_jail_pop_data <- incarceration_trends %>%
    select(
      year,
      state,
      female_jail_pop,
      male_jail_pop,
      county_name,
    ) 
View(get_king_county_jail_pop_data)
This code allowed me to compare the amount of males in jail, and females in jail
within King County, WA between the years 1970 to 2018. From this data we can visibly see that men have a much higher incarceration rate than females. This raises the question of are more males being wrongfully accussed of a crime than females, or are they just simply committing more crimes. It raises an important question. 


plot(plot_jail_pop_for_us(get_year_jail_pop))  

Between the years 1970 and 2018 we have seen a general increase in total incarcerated population
within the United States. We can see that the lowest that the incarcerated population has been was in the 
1970s and the highest was in the 2010s. We can conclude that there has been a significant increase in 
incarcerated populations but why has there been such a dramatic increase in jail population? 
  
plot(plot_jail_pop_by_states(c("WA", "NC", "CA", "NY", "FL")))

In this line graph we are looking at the total incarceration population from 1970-2018
in five states: California, Florida, North Carolina, New York, and Washington. From
this chart, we can observe a general increase in all five states, except for New York when around
2000 they start to see a general decrease in total numbers. California has had the highest amounts
of incarceration, but that is not surprising due to their population being the largest. 
A question to ask from this chart is was caused New Yorks incarceration decline around
the 2000s? 

plot(plot_king_county_data_1970_to_2018())

In this chart we are looking at King County, Washingtons total jail population by gender 
from 1970-2018. In this chart we can see an increase in jail population for both genders until 
the 2000s, however, there has been a decline in total numbers since. Another thing we can 
observe is how the population has a significanlty higher incarcerated male population
compared to females. Why do men have a much higher rate of incarceration? 


 coord_map()
 
This map is showing the top ten most populated states in the United States, and there
ratio of population in jail compared to the total population. From this map is it visible 
that states with the highest populations, for example, California, do not always have the 
highest percentage of amount of people in jail. For example, Georgia, has the highest
percentage of people in jail even though it is not the highest populated states. Why
do some states have a higher ratio of imprisoned people than other states?
_All the very best with your work!_

